---
title: 'Brain Interaction Networks'
author: 'Jon Thompson'
date: '`r Sys.Date()`'
output: "html_notebook"
output_dir: "./docs"
format: 
    html:
        code-fold: true 
---

<!-- render: `rmarkdown::render("analysis/analysis.Rmd", output_format = "html_notebook", output_dir="docs", output_file=I("index.html"))` -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo=FALSE,
    message=FALSE,
    warning=FALSE
)
```

```{r}
options(
    stringsAsFactors=F,
    use='parirwise.complete.obs',
    warn=1,
    verbose=F,
    mc.cores=5
)

randomSeed=12345
set.seed(randomSeed)
```

```{r}
library('here')
library("ggplot2")
library("magrittr")
library('data.table')
library('openxlsx')
library('rje')
# https://github.com/christophergandrud/networkD3
# how to control node colour: https://stackoverflow.com/a/35371131
library("networkD3")
library("rmarkdown")
```

```{r}
source(here('my-r-utils', 'utility.R'))
```

```{r}
xls_sheets_to_list_dt <- function(excel_file, skip_first_sheet=T) {
	sheet_names <- openxlsx::getSheetNames(excel_file)
	if (skip_first_sheet) sheet_names = sheet_names[2:length(sheet_names)]
	list_dt = lapply(
		X=sheet_names, 
		FUN=function(sheet) {
			openxlsx::read.xlsx(
			excel_file,
			sheet = sheet,
			startRow = 1
			) %>% data.table::as.data.table()
		}
	)
	names(list_dt) = sheet_names
	return(list_dt)
}
```


```{r}
# suppl data 9: long binary lists of interactors per experiment, including combined networks across time points and index proteins

file_suppl_data_9 = here("data", "Tables", "SCZ_Data_S9.xlsx")
list_dt_suppl_data_9 = xls_sheets_to_list_dt(file_suppl_data_9)
```

# analysis 

These analyses focus on overlaps 

1. between replicates of the same index protein at different weeks
2. between the networks that have been combined by index across replicates

## plot networks as graphs

```{r}
# prepare lists with different IP replicates of the same index protein
# proteins for which multiple IP replicates exist
proteins_replicated = c("CACNA1C", "HCN1", "SYNGAP1")
```
```{r}
list_nw_replicated = lapply(proteins_replicated, function(protein) {
	nw = grep(protein, names(list_dt_suppl_data_9), value=T)
	return(grep("combined", nw, invert=T, value=T))
})

names(list_nw_replicated) = proteins_replicated
```

```{r}
list_list_nw_replicated_proteins = lapply(
	list_nw_replicated, 
	function(replicates) {
		list_out = lapply(
			replicates, 
			function(replicate) list_dt_suppl_data_9[[replicate]][IsInteractor==TRUE, Gene]
		)
		names(list_out) = replicates
		return(list_out)
	}
)

```

```{r}
# prepare lists of networks combined across replicates for each index protein
nw_combined_sheets = c(
	"CACNA1C_combined",
	"HCN1_combined",
	"RIMS1_combined",
	"SYNGAP1_combined",
	"CUL3_wk7",
	"TCF4_wk1"
)
```
```{r}
list_proteins_combined_by_index = lapply(
	X = nw_combined_sheets,
	FUN = function(sheet) {
		list_dt_suppl_data_9[[sheet]][IsInteractor==T, Gene]
	}
)
names(list_proteins_combined_by_index) = nw_combined_sheets
```

### different replicates of the same index protein


```{r}
# prepare links and nodes
# source and target entries refer to row of dt_nodes_comb, but are 0-based (so -1)

list_list_dt_replicated_proteins = lapply(
    names(list_list_nw_replicated_proteins), 
    function(index_protein) {
		lapply(
			names(list_list_nw_replicated_proteins[[index_protein]]),
			function(replicate) {
				data.table(
					name = list_list_nw_replicated_proteins[[index_protein]][[replicate]]
				)
			}
		) %>% data.table::rbindlist() -> dt_nodes
		# add replicates as nodes
		dt_nodes = data.table::rbindlist(
			list(
				dt_nodes, 
				data.table(
					name=names(list_list_nw_replicated_proteins[[index_protein]])
				)
			)
		)
		dt_nodes = unique(dt_nodes)
		dt_nodes$id = 0:(nrow(dt_nodes)-1)
		# links
		dt_links = lapply(
			names(list_list_nw_replicated_proteins[[index_protein]]),
			function(replicate) {
				# list_list_nw_replicated_proteins[[index_protein]][[replicate]] %>% 
				# 	{
				data.table(
					source = which(dt_nodes$name == replicate)-1,
					target = sapply(
						list_list_nw_replicated_proteins[[index_protein]][[replicate]],
						function(interactor) which(dt_nodes$name == interactor)-1
					)
				)
					# }
			}
		) %>% data.table::rbindlist() -> dt_links

		# add node degree to nodes
		node_degree_interactors = dt_links[, .N, by=target]
		dt_nodes$degree = node_degree_interactors$N[match(dt_nodes$id, node_degree_interactors$target)]
		node_degree_index = dt_links[, .N, by=source]
		dt_nodes$degree[match(node_degree_index$source,dt_nodes$id)] = node_degree_index$N
		
		# add group as unique intersections
		dt_nodes$group = sapply(
			dt_nodes$id, 
			function(x) {
				dt_links$source[which(dt_links$target==x)] %>% 
				sort -> source_node_ids
				dt_nodes$name[source_node_ids+1] %>%
				paste0(collapse="&")
				}
			)
		return(
			list(
				"nodes"=dt_nodes,
				"links"=dt_links
			)
		)
	}
)

names(list_list_dt_replicated_proteins) = names(list_list_nw_replicated_proteins)
```

#### CACNA1C
```{r}

networkD3::forceNetwork(
	Links = list_list_dt_replicated_proteins$CACNA1C$links, 
	Nodes = list_list_dt_replicated_proteins$CACNA1C$nodes, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)')
```


#### HCN1
```{r}

networkD3::forceNetwork(
	Links = list_list_dt_replicated_proteins$HCN1$links, 
	Nodes = list_list_dt_replicated_proteins$HCN1$nodes, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)'
)
```

#### SYNGAP1
```{r}

networkD3::forceNetwork(
	Nodes = list_list_dt_replicated_proteins$SYNGAP1$nodes,
	Links = list_list_dt_replicated_proteins$SYNGAP1$links, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)')
```


### combined networks

```{r}
# prepare links and nodes
dt_nodes_comb = data.table::data.table(
	name = c(
		unlist(list_proteins_combined_by_index, use.names=F),
		gsub("_.*", "", nw_combined_sheets)
		)
)

dt_nodes_comb = unique(dt_nodes_comb)
dt_nodes_comb$id = 0:(nrow(dt_nodes_comb)-1)

# links
# source and target entries refer to row of dt_nodes_comb, but are 0-based (so -1)
dt_links = lapply(
	names(list_proteins_combined_by_index),
	function(index_protein_comb) {
		data.table(
			source = which(dt_nodes_comb$name == gsub("_.*","",index_protein_comb))-1,
			target = sapply(
				list_proteins_combined_by_index[[index_protein_comb]],
				function(interactor) which(dt_nodes_comb$name == interactor)-1
			)
		)
	}
) %>% data.table::rbindlist() -> dt_links_comb

# add node degree to nodes
node_degree_interactors = dt_links_comb[, .N, by=target]
dt_nodes_comb$degree = node_degree_interactors$N[match(dt_nodes_comb$id, node_degree_interactors$target)]
node_degree_index = dt_links_comb[, .N, by=source]
dt_nodes_comb$degree[match(node_degree_index$source,dt_nodes_comb$id)] = node_degree_index$N

# add group as unique intersections
dt_nodes_comb$group = sapply(
	dt_nodes_comb$id, 
	function(x) {
		dt_links_comb$source[which(dt_links_comb$target==x)] %>% 
		sort -> source_node_ids
		dt_nodes_comb$name[source_node_ids+1] %>%
		paste0(collapse="&")
	}
)

```

```{r}

networkD3::forceNetwork(
	Nodes = dt_nodes_comb,
	Links = dt_links_comb, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)')
```


## How do overlaps change under different LFC fold-change cutoffs?

The criteria for including a protein as interactor are generous:
".. defining proteins with log2 FC > 0 and FDR ≤ 0.1 as the significant interactors of the index  protein.” ([Hsu et al., 2022, p. 8])"


```{r}
# load data
file_suppl_data_6 = here("data", "Tables", "SCZ_Data_S6.xlsx")
list_dt_suppl_data_6 = xls_sheets_to_list_dt(file_suppl_data_6)
```


```{r}
# only keep log(FC)>0
list_dt_suppl_data_6_pos = lapply(list_dt_suppl_data_6, function(dt) dt[logFC>0])
```


```{r}
# add 'experiment' and 'index' columns
for (experiment in names(list_dt_suppl_data_6_pos)) {
	list_dt_suppl_data_6_pos[[experiment]]$experiment = experiment
	list_dt_suppl_data_6_pos[[experiment]]$index = gsub("_.*", "", experiment)
}
```

```{r}
dt_suppl_data_6_comb = data.table::rbindlist(list_dt_suppl_data_6_pos)
```

### how does fold-change FDR cutoff affect the total number of interactors?

```{r}
# define plot function

plot_count_by_cutoff = function(
	dt,
	ylab="n_interactors",
	cutoff_variable="FDR",
	cutoff_values=seq(0,0.1,0.001)
	) {

	dt_cutoff = data.table::data.table(
		cutoff = cutoff_values,
		observed = sapply(
			cutoff_values,
			function(cutoff) dt[dt[[cutoff_variable]] < cutoff, .N]
		)
	)
	dt_cutoff$expected = seq(0, max(dt_cutoff$observed), length.out=nrow(dt_cutoff))

	dt_long = data.table::melt(
		dt_cutoff, 
		id.vars="cutoff"
	)

	# make a line plot with two lines: observed and expected
	optimal_cutoff = dt_cutoff$cutoff[which.max(dt_cutoff$observed-dt_cutoff$expected)]

	print(paste0("optimal cutoff:", optimal_cutoff))
	
	ggplot(
		dt_long, 
		aes(x=cutoff, y=value, color=variable)
	) +
	geom_line() +
	theme_bw() +
	xlab(paste0(cutoff_variable, " cutoff")) +
	ylab(ylab) +
	theme(
		axis.text.x = element_text(angle=90, hjust=1)
	) +
	geom_vline(
		xintercept = optimal_cutoff,
		color = "red",
		linetype = "dashed"
	)
}
```

select cutoff by 

1. adding together all experiments.
2. taking point where distance between observed and expected combined network size (KS-stat) is maximized, where 'expected' is on the assumption that interactor FDR p-values are uniformly distributed.

> [!info] Note that we should probably consider re-computing FDR values! However we are using FDR values simply as a measure of association, not for inference)

```{r}
plot_count_by_cutoff(
	dt = dt_suppl_data_6_comb,
	ylab="n_interactors",
	cutoff_variable="FDR",
	cutoff_values=seq(0,0.1,0.001)
	)
```


NB: we do not actually use this cutoff to filter networks

### how does FDR cutoff affects the mean degree of an interactor?

In this section, we look at the relationship between the LFC FDR cutoff for interactors and the number of index proteins for which they come up

```{r}

index_proteins = names(list_dt_suppl_data_6_pos) %>% 
	gsub("_.*", "", .) %>% unique

list_experiments_by_index = lapply(
	index_proteins,
	function(index) {
		names(list_dt_suppl_data_6_pos) %>% 
		grep(index, ., value=T)
	}
)
names(list_experiments_by_index) = index_proteins
```

```{r}
list_dt_combined_by_index = lapply(
	names(list_experiments_by_index),
	function(index_protein) {
		data.table::rbindlist(
			list_dt_suppl_data_6_pos[list_experiments_by_index[[index_protein]]]
		)
	})
dt_combined_by_index = data.table::rbindlist(list_dt_combined_by_index)
```


```{r}
# define plot function
plot_degree_by_gene_by_cutoff = function(
	dt,
	ylab="mean_degree",
	cutoff_variable="FDR",
	cutoff_values=seq(0.001,0.1,0.001)
	) {
	
	sapply(cutoff_values, function(cutoff) {
		dt_cutoff = dt[dt[[cutoff_variable]] < cutoff]
		dt_cutoff = dt_cutoff[!duplicated(dt_cutoff[, .(gene, index)])]
		mean(dt_cutoff[, .(.N), by=gene]$N)
	}) -> mean_degree_by_cutoff

	mean_degree_by_cutoff[is.na(mean_degree_by_cutoff)] = 0

	
	dt_plot = data.table::data.table(
		cutoff = cutoff_values,
		mean_degree=mean_degree_by_cutoff
	)

	ggplot(
		dt_plot, 
		aes(x=cutoff, y=mean_degree)
	) +
	geom_line() +
	theme_bw() +
	xlab(paste0(cutoff_variable, " cutoff")) +
	ylab(ylab) +
	theme(
		axis.text.x = element_text(angle=90, hjust=1)
	) 
}
```

```{r}
plot_degree_by_gene_by_cutoff(
	dt = dt_combined_by_index,
	ylab="mean_degree",
	cutoff_variable="FDR",
	cutoff_values=seq(0.00,1,0.01)
	)
```

The mean degree of interactors rises quickly initially, then slower as we raise FDR cutoff.

This suggests that the overlap is not noise, since reoccurring interactors tend to have highly significant FC FDR p-values.

Next, we look at the relationship between an interactor's mean (across indices) minimum (across replicates) FDR, and the number of indexes for which it comes up. 
I.e.: do recurring interactors tend to have low FDR p-values?

We look at the proteins with non-negative logFC and FDR p-values < 0.1.
```{r}

```{r}
# remove interactors with FDR >= 0.1
dt_cutoff = dt_combined_by_index[FDR < 0.1]
# mark interactor genes that appear in multiple replicates of the same index
dt_cutoff[,c("duplicated"):=duplicated(dt_cutoff[, .(gene, index)])]
# add a new column with the lowest FDR of each gene within each index (across replicates)
dt_cutoff[, c("min_FDR"):=min(FDR), by=.(gene, index)]
# leave only one replicate of each gene per index
dt_cutoff = dt_cutoff[duplicated==F]
# count how many indices each gene appears in
dt_cutoff[, N:=.N, by=gene]
# take the mean min FDR across indices
dt_cutoff[, mean_min_FDR:=mean(min_FDR), by=gene]
# leave one row per gene
dt_cutoff = dt_cutoff[!duplicated(gene)]
# r$> any(duplicated(dt_cutoff$gene))
# [1] FALSE
```

```{r}
# we want to degree vs mean minimum FDR across indices
ggplot(
	dt_cutoff, 
	aes(x=mean_min_FDR, y=N)
) + 
geom_point() +
theme_bw() +
	xlab("mean minimum FDR across indices") +
	ylab("number of indices") +
	theme(
		axis.text.x = element_text(angle=90, hjust=1)
	) 

```

We see that the interactors with an FDR < 0.1 that occur across most indices tend to have a low (highly statistically significant) FDR.

This may suggest that there is a set of interactors that are robustly shared across the index proteins.

What are the hub interactors that are shared between 5 out of 6 index proteins?

```{r}
# show the genes with N>=5 ordered by min_mean_FDR
dt_cutoff[N>=5][order(mean_min_FDR)][, .(gene, mean_min_FDR)]
```

None of these are index genes.

What are the interactors that are shared between 4 out of 6 index proteins?

```{r}
# show the genes with N>=5 ordered by min_mean_FDR
dt_cutoff[N>=4][order(mean_min_FDR)][, .(gene, mean_min_FDR)]
```

## PRS analyses

```{r}
# prepare data for PRS analyses of overlaps

overlap_all_pairs = function(
	list_vec,
	overlap_func = function(x) {x}
	) {
	# takes the powerset of list_vec and performs overlap_func on each subset (of sets)
	#
	# @param list_vec: a list of vectors
	# @param overlap_func: some function to apply to each subset of list_vec
	# @return: named list of vectors of length 2^length(list_vec) - 1
	combs = rje::powerSet(names(list_vec))
	# throw away empty set
	combs = combs[2:length(combs)]
	overlaps = sapply(
		X = combs, 
		FUN = function(comb) { 
			# comb is a character vector
			list_vec = list_vec[comb]
			overlap_func(list_vec)
		}
	)
	names(overlaps) = sapply(combs, function(x) paste0(x, collapse="&"))
	return(overlaps)
}
```

```{r}
intersect_reduce = function(list_vec) Reduce(f=intersect, x=list_vec)
```
```{r}
overlap_length = function(list_vec) Reduce(f=intersect, x=list_vec) %>% length()
```

```{r}
overlap_all_pairs_proteins_combined_by_index = overlap_all_pairs(
	list_vec=list_proteins_combined_by_index,
	overlap_func = intersect_reduce
	) 
``` 

next steps (done outside this script): investigate PRS of overlaps

* interactors with degree = 5
* every subset of interactors with degree = 4
* every subset of interactors with degree = 3
* every subset of interactors with degree = 2
* every subset of interactors with degree = 1
