---
title: ''
author: 'Jon Thompson'
date: '`r Sys.Date()`'
output: "html_notebook"
format: 
    html:
        code-fold: true 
---

<!-- to render, do `rmarkdown::render("analysis/analysis.Rmd", output_format = "html_notebook")` -->

# setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo=FALSE,
    message=FALSE,
    warning=FALSE
)
```

```{r}
options(
    stringsAsFactors=F,
    use='parirwise.complete.obs',
    warn=1,
    verbose=F,
    mc.cores=5
)

randomSeed=12345
set.seed(randomSeed)
```

## libraries

```{r}
library('here')
library("ggplot2")
library("magrittr")
library('data.table')
library('openxlsx')
library('rje')
# https://github.com/christophergandrud/networkD3
library("networkD3")
library("rmarkdown")
```

## utility functions

```{r}
source(here('my-r-utils', 'utility.R'))
```

## load Hsu et al preprint supplementary data

```{r}
xls_sheets_to_list_dt <- function(excel_file, skip_first_sheet=T) {
	sheet_names <- openxlsx::getSheetNames(excel_file)
	if (skip_first_sheet) sheet_names = sheet_names[2:length(sheet_names)]
	list_dt = lapply(
		X=sheet_names, 
		FUN=function(sheet) {
			openxlsx::read.xlsx(
			excel_file,
			sheet = sheet,
			startRow = 1
			) %>% data.table::as.data.table()
		}
	)
	names(list_dt) = sheet_names
	return(list_dt)
}
```

* suppl data 9: long binary lists of interactors per experiment, including combined networks across time points and index proteins

```{r}
file_suppl_data_9 = here("data", "Tables", "SCZ_Data_S9.xlsx")
list_dt_suppl_data_9 = xls_sheets_to_list_dt(file_suppl_data_9)
```

# analysis 

These analyses focus on overlaps

## plot network and overlaps as force directed graphs

prepare lists with different IP replicates of the same index protein

```{r}
# proteins for which multiple IP replicates exist
proteins_replicated = c("CACNA1C", "HCN1", "SYNGAP1")
```
```{r}
list_nw_replicated = lapply(proteins_replicated, function(protein) {
	nw = grep(protein, names(list_dt_suppl_data_9), value=T)
	return(grep("combined", nw, invert=T, value=T))
})

names(list_nw_replicated) = proteins_replicated
```

```{r}
list_list_nw_replicated_proteins = lapply(
	list_nw_replicated, 
	function(replicates) {
		list_out = lapply(
			replicates, 
			function(replicate) list_dt_suppl_data_9[[replicate]][IsInteractor==TRUE, Gene]
		)
		names(list_out) = replicates
		return(list_out)
	}
)

```

prepare lists of networks combined across replicates for each index protein
```{r}
nw_combined_sheets = c(
	"CACNA1C_combined",
	"HCN1_combined",
	"RIMS1_combined",
	"SYNGAP1_combined",
	"CUL3_wk7",
	"TCF4_wk1"
)
```
```{r}
list_proteins_combined_by_index = lapply(
	X = nw_combined_sheets,
	FUN = function(sheet) {
		list_dt_suppl_data_9[[sheet]][IsInteractor==T, Gene]
	}
)
names(list_proteins_combined_by_index) = nw_combined_sheets
```

### plot overlaps between different replicates of the same index protein

prepare links and nodes
```{r}
# source and target entries refer to row of dt_nodes_comb, but are 0-based (so -1)

list_list_dt_replicated_proteins = lapply(
    names(list_list_nw_replicated_proteins), 
    function(index_protein) {
		lapply(
			names(list_list_nw_replicated_proteins[[index_protein]]),
			function(replicate) {
				data.table(
					name = list_list_nw_replicated_proteins[[index_protein]][[replicate]]
				)
			}
		) %>% data.table::rbindlist() -> dt_nodes
		# add replicates as nodes
		dt_nodes = data.table::rbindlist(
			list(
				dt_nodes, 
				data.table(
					name=names(list_list_nw_replicated_proteins[[index_protein]])
				)
			)
		)
		dt_nodes = unique(dt_nodes)
		dt_nodes$id = 0:(nrow(dt_nodes)-1)
		# links
		dt_links = lapply(
			names(list_list_nw_replicated_proteins[[index_protein]]),
			function(replicate) {
				# list_list_nw_replicated_proteins[[index_protein]][[replicate]] %>% 
				# 	{
				data.table(
					source = which(dt_nodes$name == replicate)-1,
					target = sapply(
						list_list_nw_replicated_proteins[[index_protein]][[replicate]],
						function(interactor) which(dt_nodes$name == interactor)-1
					)
				)
					# }
			}
		) %>% data.table::rbindlist() -> dt_links

		# add node degree to nodes
		node_degree_interactors = dt_links[, .N, by=target]
		dt_nodes$degree = node_degree_interactors$N[match(dt_nodes$id, node_degree_interactors$target)]
		node_degree_index = dt_links[, .N, by=source]
		dt_nodes$degree[match(node_degree_index$source,dt_nodes$id)] = node_degree_index$N
		
		# add group as unique intersections
		dt_nodes$group = sapply(
			dt_nodes$id, 
			function(x) {
				dt_links$source[which(dt_links$target==x)] %>% 
				sort -> source_node_ids
				dt_nodes$name[source_node_ids+1] %>%
				paste0(collapse="&")
				}
			)
		return(
			list(
				"nodes"=dt_nodes,
				"links"=dt_links
			)
		)
	}
)

names(list_list_dt_replicated_proteins) = names(list_list_nw_replicated_proteins)
```

plot different replicates of CACNA1C
```{r}

networkD3::forceNetwork(
	Links = list_list_dt_replicated_proteins$CACNA1C$links, 
	Nodes = list_list_dt_replicated_proteins$CACNA1C$nodes, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)')
```

plot different replicates of HCN1
```{r}

networkD3::forceNetwork(
	Links = list_list_dt_replicated_proteins$HCN1$links, 
	Nodes = list_list_dt_replicated_proteins$HCN1$nodes, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)'
)
```

plot different replicates of SYNGAP1
```{r}

networkD3::forceNetwork(
	Nodes = list_list_dt_replicated_proteins$SYNGAP1$nodes,
	Links = list_list_dt_replicated_proteins$SYNGAP1$links, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)')
```

### plot overlaps between combined networks

prepare links and nodes
```{r}
dt_nodes_comb = data.table::data.table(
	name = c(
		unlist(list_proteins_combined_by_index, use.names=F),
		gsub("_.*", "", nw_combined_sheets)
		)
)

dt_nodes_comb = unique(dt_nodes_comb)
dt_nodes_comb$id = 0:(nrow(dt_nodes_comb)-1)

# links
# source and target entries refer to row of dt_nodes_comb, but are 0-based (so -1)
dt_links = lapply(
	names(list_proteins_combined_by_index),
	function(index_protein_comb) {
		data.table(
			source = which(dt_nodes_comb$name == gsub("_.*","",index_protein_comb))-1,
			target = sapply(
				list_proteins_combined_by_index[[index_protein_comb]],
				function(interactor) which(dt_nodes_comb$name == interactor)-1
			)
		)
	}
) %>% data.table::rbindlist() -> dt_links_comb

# add node degree to nodes
node_degree_interactors = dt_links_comb[, .N, by=target]
dt_nodes_comb$degree = node_degree_interactors$N[match(dt_nodes_comb$id, node_degree_interactors$target)]
node_degree_index = dt_links_comb[, .N, by=source]
dt_nodes_comb$degree[match(node_degree_index$source,dt_nodes_comb$id)] = node_degree_index$N

# add group as unique intersections
dt_nodes_comb$group = sapply(
	dt_nodes_comb$id, 
	function(x) {
		dt_links_comb$source[which(dt_links_comb$target==x)] %>% 
		sort -> source_node_ids
		dt_nodes_comb$name[source_node_ids+1] %>%
		paste0(collapse="&")
	}
)

```

```{r}

networkD3::forceNetwork(
	Nodes = dt_nodes_comb,
	Links = dt_links_comb, 
	Source = "source", 
	Target = "target",
	# Value = "degree", 
	Nodesize = "degree",
	NodeID = "name",
	Group = "group", 
	fontSize=15,
	opacity=0.8, 
	zoom = T,
    clickAction = 'alert(d.name)')
```

## investigate effects of different fold-change cutoffs

Given the large differences in the number of detected interactors when using the same index at the same week timepoint, plot the network size as a function of FC cut-off to check for whether there a 'shoulder' in the distribution of network size.

From the manuscript: ".. defining proteins with log2 FC > 0 and FDR ≤ 0.1 as the significant interactors of the index  protein.” ([Hsu et al., 2022, p. 8])"

load data
```{r}
file_suppl_data_6 = here("data", "Tables", "SCZ_Data_S6.xlsx")
list_dt_suppl_data_6 = xls_sheets_to_list_dt(file_suppl_data_6)
```


```{r}
# only keep log(FC)>0
list_dt_suppl_data_6_pos = lapply(list_dt_suppl_data_6, function(dt) dt[logFC>0])
```

add 'experiment' and 'index' columns
```{r}
for (experiment in names(list_dt_suppl_data_6_pos)) {
	list_dt_suppl_data_6_pos[[experiment]]$experiment = experiment
	list_dt_suppl_data_6_pos[[experiment]]$index = gsub("_.*", "", experiment)
}
```

```{r}
dt_suppl_data_6_comb = data.table::rbindlist(list_dt_suppl_data_6_pos)
```

### how does fold-change FDR cutoff affect total number of interactors?

```{r}
# define plot function

plot_count_by_cutoff = function(
	dt,
	ylab="n_interactors",
	cutoff_variable="FDR",
	cutoff_values=seq(0,0.1,0.001)
	) {

	dt_cutoff = data.table::data.table(
		cutoff = cutoff_values,
		observed = sapply(
			cutoff_values,
			function(cutoff) dt[dt[[cutoff_variable]] < cutoff, .N]
		)
	)
	dt_cutoff$expected = seq(0, max(dt_cutoff$observed), length.out=nrow(dt_cutoff))

	dt_long = data.table::melt(
		dt_cutoff, 
		id.vars="cutoff"
	)

	# make a line plot with two lines: observed and expected
	optimal_cutoff = dt_cutoff$cutoff[which.max(dt_cutoff$observed-dt_cutoff$expected)]

	print(paste0("optimal cutoff:", optimal_cutoff))
	
	ggplot(
		dt_long, 
		aes(x=cutoff, y=value, color=variable)
	) +
	geom_line() +
	theme_bw() +
	xlab(paste0(cutoff_variable, " cutoff")) +
	ylab(ylab) +
	theme(
		axis.text.x = element_text(angle=90, hjust=1)
	) +
	geom_vline(
		xintercept = optimal_cutoff,
		color = "red",
		linetype = "dashed"
	)
}
```

* select cutoff by 
1. adding together all experiments
2. taking point where distance between observed and expected combined network size (KS-stat) is maximized, where expected assumes a uniform distribution of interactor FDR p-values 

> [!info] Note that we should probably consider re-computing FDR values! However we are using FDR values simply as a measure of association, not for inference)


```{r}
plot_count_by_cutoff(
	dt = dt_suppl_data_6_comb,
	ylab="n_interactors",
	cutoff_variable="FDR",
	cutoff_values=seq(0,0.1,0.001)
	)
```

### how does FDR cutoff affects the mean degree of an interactor?

NB: do not filter networks with a more stringent FDR cutoff using the approach of the previous section. This is a separate analysis to see how FDR affects the index-degree of interactors (i.e. the number of different index proteins for which the same interactor comes up)

```{r}

index_proteins = names(list_dt_suppl_data_6_pos) %>% 
	gsub("_.*", "", .) %>% unique

list_experiments_by_index = lapply(
	index_proteins,
	function(index) {
		names(list_dt_suppl_data_6_pos) %>% 
		grep(index, ., value=T)
	}
)
names(list_experiments_by_index) = index_proteins
```

```{r}
list_dt_combined_by_index = lapply(
	names(list_experiments_by_index),
	function(index_protein) {
		data.table::rbindlist(
			list_dt_suppl_data_6_pos[list_experiments_by_index[[index_protein]]]
		)
	})
dt_combined_by_index = data.table::rbindlist(list_dt_combined_by_index)
```

define plot function
```{r}
plot_degree_by_gene_by_cutoff = function(
	dt,
	ylab="mean_degree",
	cutoff_variable="FDR",
	cutoff_values=seq(0.001,0.1,0.001)
	) {
	
	sapply(cutoff_values, function(cutoff) {
		dt_cutoff = dt[dt[[cutoff_variable]] < cutoff]
		dt_cutoff = dt_cutoff[!duplicated(dt_cutoff[, .(gene, index)])]
		mean(dt_cutoff[, .(.N), by=gene]$N)
	}) -> mean_degree_by_cutoff

	mean_degree_by_cutoff[is.na(mean_degree_by_cutoff)] = 0

	
	dt_plot = data.table::data.table(
		cutoff = cutoff_values,
		mean_degree=mean_degree_by_cutoff
	)

	ggplot(
		dt_plot, 
		aes(x=cutoff, y=mean_degree)
	) +
	geom_line() +
	theme_bw() +
	xlab(paste0(cutoff_variable, " cutoff")) +
	ylab(ylab) +
	theme(
		axis.text.x = element_text(angle=90, hjust=1)
	) 
}
```

```{r}
plot_degree_by_gene_by_cutoff(
	dt = dt_combined_by_index,
	ylab="mean_degree",
	cutoff_variable="FDR",
	cutoff_values=seq(0.00,1,0.01)
	)
```

The mean degree of interactors rises quickly initially, then slower as we raise FDR cutoff
This suggests that the overlap is not noise, since reoccurring interactors tend to have highly significant FC FDR p-values

#### what is the relationship between an interactor's mean (across indices) minimum (across replicates) FDR, and the number of indexes for which it comes up?

```{r}
# remove interactors with FDR >= 0.1
dt_cutoff = dt_combined_by_index[FDR < 0.1]
# mark interactor genes that appear in multiple replicates of the same index
dt_cutoff[,c("duplicated"):=duplicated(dt_cutoff[, .(gene, index)])]
# add a new column with the lowest FDR of each gene within each index (across replicates)
dt_cutoff[, c("min_FDR"):=min(FDR), by=.(gene, index)]
# leave only one replicate of each gene per index
dt_cutoff = dt_cutoff[duplicated==F]
# count how many indices each gene appears in
dt_cutoff[, N:=.N, by=gene]
# take the mean min FDR across indices
dt_cutoff[, mean_min_FDR:=mean(min_FDR), by=gene]
# leave one row per gene
dt_cutoff = dt_cutoff[!duplicated(gene)]
# r$> any(duplicated(dt_cutoff$gene))
# [1] FALSE
```

```{r}
# we want to degree vs mean minimum FDR across indices
ggplot(
	dt_cutoff, 
	aes(x=mean_min_FDR, y=N)
) + 
geom_point() +
theme_bw() +
	xlab("mean minimum FDR across indices") +
	ylab("number of indices") +
	theme(
		axis.text.x = element_text(angle=90, hjust=1)
	) 

```

When we look at the proteins with non-negative logFC and FDR p-values < 0.1, we see that the interactors with an FDR < 0.1 that occur across most indices tend to have a low (highly statistically significant) FDR.

This may suggest that there is a set of interactors that are robustly shared across the index proteins.

What are these hub interactors?

```{r}
# show the genes with N>=5 ordered by min_mean_FDR
dt_cutoff[N>=5][order(mean_min_FDR)][, .(gene, mean_min_FDR)]
```

None of the index genes are hub genes.

## PRS analyses

prepare data for PRS analyses of overlaps

```{r}

overlap_all_pairs = function(
	list_vec,
	overlap_func = function(x) {x}
	) {
	# takes the powerset of list_vec and performs overlap_func on each subset (of sets)
	#
	# @param list_vec: a list of vectors
	# @param overlap_func: some function to apply to each subset of list_vec
	# @return: named list of vectors of length 2^length(list_vec) - 1
	combs = rje::powerSet(names(list_vec))
	# throw away empty set
	combs = combs[2:length(combs)]
	overlaps = sapply(
		X = combs, 
		FUN = function(comb) { 
			# comb is a character vector
			list_vec = list_vec[comb]
			overlap_func(list_vec)
		}
	)
	names(overlaps) = sapply(combs, function(x) paste0(x, collapse="&"))
	return(overlaps)
}
```

```{r}
intersect_reduce = function(list_vec) Reduce(f=intersect, x=list_vec)
```
```{r}
overlap_length = function(list_vec) Reduce(f=intersect, x=list_vec) %>% length()
```

```{r}
overlap_all_pairs_proteins_combined_by_index = overlap_all_pairs(
	list_vec=list_proteins_combined_by_index,
	overlap_func = intersect_reduce
	) 
``` 

next steps (done outside this script): 
* investigate PRS of overlaps
	* interactors with degree = 5
	* every subset of interactors with degree = 4
	* every subset of interactors with degree = 3
	* every subset of interactors with degree = 2
	* every subset of interactors with degree = 1
