---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#####################################################
#################### DEPRECATED #####################
#####################################################

```{r}
list_overlap_all_pairs_nw_replicated_proteins = lapply(
	list_list_nw_replicated_proteins, 
	function(list_vec) overlap_all_pairs(list_vec, overlap_func=overlap_length)
)
```
```{r}
list_p_upset_wk = lapply(
	1:length(list_list_nw_replicated_proteins),
	function(i) {
		UpSetR::upset(
			# need to use fromExpression input to get all combinations!
			fromExpression(list_overlap_all_pairs_nw_replicated_proteins[[i]]),
			fromList(list_list_nw_replicated_proteins[[i]]),
			sets=names(list_list_nw_replicated_proteins[[i]]), # required by keep.order=T
			keep.order=T, # requires the 'sets' argument
			set_size.show=F
			)
	}
)
```

```{r}
list_p_upset_wk
```

###### comments on overlap between replicates

TODO

##### comments on overlap between index interactors
* every set consists of <40% unique proteins
* HCN1 and CACNA1C share around 25% of their proteins
* 51 proteins shared between all 4 networks


#### overlaps between protein sets combined across weeks
```{r}
UpSetR::upset(
	# fromExpression(overlap_all_pairs_proteins_combined_by_index),
	fromList(list_proteins_combined_by_index),
	order.by = "freq" # default
	# group.by = "sets" #  can group by set to make it easier to read # no, this duplicates them
)
```

### euler diagramts of overlaps between combined IPs from different index proteins

failed due to euler not being able to find a fit!


```{r}
nw_combined_overlaps = overlap_all_pairs(list_proteins_combined_by_index)
```

```{r}
nw_combined_overlaps_euler_fit = eulerr::euler(
	combinations = nw_combined_overlaps,
	input="union",
	# ellipse is easier to fit
	shape="ellipse"
)

```

check goodness of fit
```{r}
eulerr::error_plot(nw_combined_overlaps_euler_fit)
```

```{r}
nw_combined_overlaps_euler_fit
```

some large
```{r}
p1 = plot(
	x=nw_combined_overlaps_euler_fit,
	quantities=T,
	main="Overlaps between combined IPs from different index proteins"
	)

p1
```


### plot overlaps between combined IPs from different index proteins (stringent)

```{r}
list_proteins_combined_stringent_by_index = lapply(
	X = nw_combined_sheets,
	FUN = function(sheet) {
		if (grepl("combined", sheet)) {
			list_dt_suppl_data_9[[sheet]][IsInteractor==T && IsStringentInt==T, Gene]
		} else {
			list_dt_suppl_data_9[[sheet]][IsInteractor==T, Gene]
		}
	}
)
names(list_proteins_combined_stringent_by_index) = nw_combined_sheets
```

```{r}
nw_combined_stringent_overlaps = overlap_all_pairs(list_proteins_combined_stringent_by_index)
```

### plot overlaps between combined IPs from different index proteins (stringent)
failed due to euler not being able to find a fit
```{r}
nw_combined_stringent_overlaps_euler_fit = eulerr::euler(
	combinations = nw_combined_stringent_overlaps,
	input="union"
)

```
```{r}
p2 = plot(
	x=nw_combined_stringent_overlaps_euler_fit,
	quantities=T,
	main="Overlaps between combined stringent IPs from different index proteins"
	)

p2
```

### overlaps between different weeks with the same index protein  
```{r}
list_nw_replicates_overlaps = lapply(
	list_list_nw_replicated_proteins,
	overlap_all_pairs
)

```

```{r}
list_nw_replicates_overlaps_euler_fit = lapply(
	list_nw_replicates_overlaps,
	function(nw_replicates_overlaps) {
		eulerr::euler(
		combinations = nw_replicates_overlaps,
		input="union",
		shape="ellipse"
	)
	}
)
```

```{r}
list_nw_replicates_overlaps_euler_fit[[1]]
```
```{r}
list_p = lapply(
	names(list_nw_replicates_overlaps_euler_fit), 
	function(protein) {
		plot(
			x=list_nw_replicates_overlaps_euler_fit[[protein]],
			quantities=T,
			main=paste0(protein, " IP replicate overlaps")
		)
	}
)

list_p
```

### plot mean jacard index as a function of protein SNP p-value cutoff

use the mean jaccard index

```{r}
# function to return the jaccard index
jaccard_index = function(x, y) {
  length(base::intersect(x, y)) / length(base::union(x, y))
}
```

```{r}
# function to compute mean of all pairwise jaccard indices
jaccard_index_all_pairs = function(list_sets) {
  # get all pairwise combinations
  comb = combn(length(list_sets), 2)
  # compute jaccard index for each pair
  jaccard_index_all_pairs = sapply(
	X = 1:ncol(comb),
	FUN = function(i) {
	  jaccard_index(
		list_sets[[comb[1, i]]], list_sets[[comb[2, i]]]
	  )
	}
  )
  # return the mean jaccard index
  return(mean(jaccard_index_all_pairs))
}
```


```{r}
vec_FDR_cutoffs = seq(0, 0.05, 0.001)

vec_jaccard_index = sapply(
	X = vec_FDR_cutoffs,
	FUN = function(FDR_cutoff) {
	  # get the list of genes that pass the FDR cutoff
	  list_genes_pass_FDR_cutoff = lapply(
		X = list_tbl_suppl_data_6,
		FUN = function(tbl) {
		  tbl$gene[tbl$FDR < FDR_cutoff]
		}
	  )
	  # compute the mean jaccard index
	  jaccard_index_all_pairs(list_genes_pass_FDR_cutoff)
	}
)
```

```{r}
dt_jaccard_plot = data.table(
	FDR_cutoff = vec_FDR_cutoffs,
	mean_jaccard_index = vec_jaccard_index
)
```

```{r}
# make a line plot of the jaccard index
ggplot(
	data = dt_jaccard_plot,
	mapping = aes(
		x = FDR_cutoff,
		y = mean_jaccard_index
	)
) +
	geom_line() +
	theme_bw() +
	theme(
		axis.text.x = element_text(angle = 90, hjust = 1)
	) + scale_x_reverse()

```